---
title: "Imaris Analysis Comparison of Cluster vs. Non-Cluster Plastic Fluorescence Metrics"
format: 
    html:
        code-fold: true
        code-tools: true
        fig-align: center
editor: visual
---

``` {r}
#| label: load-packages
#| include: false

library(ggprism)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readxl)
library(here)
library(ggsignif)
library(ggpubr)
```

```{r}
#| label: load-data
#| include: false
process_image_data <- function(image_prefix) {
  # here() tells R: "Start at the root project directory,
  # look inside 'Imaris_Data', and find this file."
  cluster_center <- read_excel(
    here(
      "Imaris_Data",
      paste0(image_prefix, "_Cluster_Plastics_GFP_CenterIntensity.xlsx")
    ),
    skip = 1
  ) %>%
    select(ID, Center_Intensity = `Intensity Center`)

  cluster_max <- read_excel(
    here(
      "Imaris_Data",
      paste0(image_prefix, "_Cluster_Plastics_GFP_MaxIntensity.xlsx")
    ),
    skip = 1
  ) %>%
    select(ID, Max_Intensity = `Intensity Max`)

  cluster_mean <- read_excel(
    here(
      "Imaris_Data",
      paste0(image_prefix, "_Cluster_Plastics_GFP_MeanIntensity.xlsx")
    ),
    skip = 1
  ) %>%
    select(ID, Mean_Intensity = `Intensity Mean`)

  cluster_voxel <- read_excel(
    here(
      "Imaris_Data",
      paste0(image_prefix, "_Cluster_Plastics_GFP_Voxel.xlsx")
    ),
    skip = 1
  ) %>%
    select(ID, Voxel = `Number of Voxels`)

  cluster_tidy <- cluster_center %>%
    left_join(cluster_max, by = "ID") %>%
    left_join(cluster_mean, by = "ID") %>%
    left_join(cluster_voxel, by = "ID") %>%
    mutate(
      Image_ID = image_prefix,
      Spot_Category = "Cluster"
    )

  # --- Same Process for Non-cluster plastics ---
  noncluster_center <- read_excel(
    here(
      "Imaris_Data",
      paste0(image_prefix, "_NonCluster_Plastics_GFP_CenterIntensity.xlsx")
    ),
    skip = 1
  ) %>%
    select(ID, Center_Intensity = `Intensity Center`)

  noncluster_max <- read_excel(
    here(
      "Imaris_Data",
      paste0(image_prefix, "_NonCluster_Plastics_GFP_MaxIntensity.xlsx")
    ),
    skip = 1
  ) %>%
    select(ID, Max_Intensity = `Intensity Max`)

  noncluster_mean <- read_excel(
    here(
      "Imaris_Data",
      paste0(image_prefix, "_NonCluster_Plastics_GFP_MeanIntensity.xlsx")
    ),
    skip = 1
  ) %>%
    select(ID, Mean_Intensity = `Intensity Mean`)

  noncluster_voxel <- read_excel(
    here(
      "Imaris_Data",
      paste0(image_prefix, "_NonCluster_Plastics_GFP_Voxel.xlsx")
    ),
    skip = 1
  ) %>%
    select(ID, Voxel = `Number of Voxels`)

  noncluster_tidy <- noncluster_center %>%
    left_join(noncluster_max, by = "ID") %>%
    left_join(noncluster_mean, by = "ID") %>%
    left_join(noncluster_voxel, by = "ID") %>%
    mutate(Image_ID = image_prefix, Spot_Category = "Non cluster")

  # # ==========================================
  # # 3. SMASH TOGETHER INTO THE MASTER DATASET
  # # ==========================================
  master_tidy_spots <- bind_rows(cluster_tidy, noncluster_tidy)
}

# Read only the ID and the raw metric from each independent file
# List all your image prefixes
image_list <- c("image1", "image14", "image28")

# Read and combine everything automatically
final_table <- map_df(image_list, process_image_data)

final_table <- final_table %>%
  mutate(Image_ID = recode(Image_ID,
    "image1"  = "Image 1",
    "image14" = "Image 14",
    "image28" = "Image 28"
  ))
```


# Imaris Fluorescence Metric Comparison 

The following is a comprehensive breakdown of fluorescence intensity metrics, including quality, mean intensity, max intensity, and voxels. The intent is to compare at two levels - cluster and non-cluster plastics, and across different images as well. 

## The Goal

The intent is to be able to visualize these metrics so that a (somewhat) universal filter can be established to help quantify plastic uptake in the CHT, while also excluding out clustered plastics that may result from macrophage uptake. 

All source code for the following analysis scripts can be found [**here**](https://github.com/GlaxoCat/R-learning/tree/main)


``` {r}
#| label: plot-cluster-plastic-voxel-distribution 
#| message: false
#| warning: false
#| fig-cap: "Voxels per spot for clustered plastics across images"
#| fig-width: 9
#| fig-height: 6

voxel_compares <- list(
  c("Image 1", "Image 14"),
  c("Image 1", "Image 28"),
  c("Image 14", "Image 28")
)
final_table %>%
  ggplot(aes(x = Image_ID, y = Voxel, fill = Image_ID, color = Image_ID)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.4, size = 1) +
  facet_wrap(~Spot_Category) +
  labs(
    title = "Plastics Distribution across Images",
    x = "Voxel",
    y = "freq"
  ) +
  theme_prism(base_size = 14) +
  scale_fill_prism(palette = "prism_light") +
  scale_color_prism(palette = "prism_light") +
  stat_compare_means(
    comparisons = voxel_compares,
    method = "t.test",
    label = "p.signif"
  )
```

If we zoom in to Image 1 - which is a *kdrl:mCherry - stab2 crispant* zebrafish injected with **50nm plastics**, the voxel comparison looks like the following:


``` {r}
#| label: plot-cluster-noncluster-voxel-distribution-image1
#| message: false
#| warning: false
#| fig-cap: "Voxels per spot for clustered vs. nonclustered plastics in Image 1"
#| fig-width: 9
#| fig-height: 6
voxel_image1_compare <- list(c("Cluster", "Non cluster"))

final_table %>%
  filter(Image_ID == "Image 1") %>%
  ggplot(aes(
    x = Spot_Category,
    y = Voxel,
    color = Spot_Category,
    fill = Spot_Category
  )) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.4, size = 1) +
  labs(
    title = "Cluster vs. NonCluster Voxel in Image 1",
    x = "Spot Category",
    y = "Voxel"
  ) +
  theme_prism(base_size = 14) +
  scale_fill_prism(palette = "prism_light") +
  scale_color_prism(palette = "prism_light") +
  stat_compare_means(
    comparisons = voxel_image1_compare,
    method = "t.test",
    label = "p.signif"
  )

```

If we put these comparisons together side by side, we then observe the voxel differences between the two:

``` {r}
#| label: plot-adjacent
#| fig-subcap:
#| - "Voxels per spot for clustered plastics across all images"
#| - "Voxels per spot for clustered vs. non-clustered within Image 1"
#| layout-ncol: 2
#| column: page
#| fig-width: 9
#| fig-height: 6

voxel_compares <- list(
  c("Image 1", "Image 14"),
  c("Image 1", "Image 28"),
  c("Image 14", "Image 28")
)
final_table %>%
  ggplot(aes(x = Image_ID, y = Voxel, fill = Image_ID, color = Image_ID)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.4, size = 1) +
  facet_wrap(~Spot_Category) +
  labs(
    title = "Plastics Distribution across Images",
    x = "Voxel",
    y = "freq"
  ) +
  theme_prism(base_size = 14) +
  scale_fill_prism(palette = "prism_light") +
  scale_color_prism(palette = "prism_light") +
  stat_compare_means(
    comparisons = voxel_compares,
    method = "t.test",
    label = "p.signif"
  )

voxel_image1_compare <- list(c("Cluster", "Non cluster"))

final_table %>%
  filter(Image_ID == "Image 1") %>%
  ggplot(aes(
    x = Spot_Category,
    y = Voxel,
    color = Spot_Category,
    fill = Spot_Category
  )) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.4, size = 1) +
  labs(
    title = "Cluster vs. NonCluster Voxel in Image 1",
    x = "Spot Category",
    y = "Voxel"
  ) +
  theme_prism(base_size = 14) +
  scale_fill_prism(palette = "prism_light") +
  scale_color_prism(palette = "prism_light") +
  stat_compare_means(
    comparisons = voxel_image1_compare,
    method = "t.test",
    label = "p.signif"
  )

```



